IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2015-01-15 paper/concrete.tex
pass over everything except related and conclusion

If the operational rules are adapted na\"ively to this new setting,
then non-interference can be violated: as we mentioned earlier,
shared mutable cells could result in violation of non-interference.
What we would like is a way of characterizing safe modifications to
the semantics which preserve non-interference.
The intention of our single heap implementation is to permit
efficient execution while \emph{conceptually maintaining isolation between
tasks} (by not allowing sharing of references between them).
This intuition of having a different (potentially more efficient)
concrete semantics that behaves like the abstract semantics
can be formalized by the following definition:

@@

\begin{theorem}[Isomorphism preserves termination sensitive non-interference]
  \label{thm:iso-tsni}
  If $L$ is isomorphic to $L'$ and $L'$ satisfies TSNI, then
  $L$ satisfies TSNI.
\end{theorem}

@@

\begin{theorem}[Weak isomorphism preserves termination insensitive non-interference]
  \label{thm:iso-tini}
  If a language $L$ is weakly isomorphic to a language $L'$, and $L'$
  satisfies TINI, then $L$ satisfies TINI.
\end{theorem}

@@

\begin{definition}[Restricted IFC language]
  \label{def:restricted}
  For a family of predicates $\mathcal P$ (one for every reduction
  rule), we call
  |restrictedLang alpha targetLang| a restricted IFC language
  if its definition is equivalent to the abstract language
  |specLang alpha targetLang|, with the following exception:
  The reduction rules are restricted
  by adding a predicate $P \in \mathcal P$ to the premise of
  all rules other than \textsc{I-noStep}.  Furthermore, the predicate $P$
  can depend only on the \textit{erased} configuration
  |erase il ic|, where |il| is the label of the first task
  in the task list and |ic| the full configuration.
\end{definition}

>>>>

If the operational rules are adapted na\"ively to this new setting,
then non-interference can be violated: as we mentioned earlier,
shared mutable cells could be used to leak sensitive information.
What we would like is a way of characterizing safe modifications to
the semantics which preserve non-interference.
The intention of our single heap implementation is to permit
efficient execution while \emph{conceptually maintaining isolation between
tasks} (by not allowing sharing of references between them).
This intuition of having a different (potentially more efficient)
concrete semantics that behaves like the abstract semantics
can be formalized by the following definition:

@@

\begin{theorem}[Isomorphism preserves TSNI]
  \label{thm:iso-tsni}
  If $L$ is isomorphic to $L'$ and $L'$ satisfies TSNI, then
  $L$ satisfies TSNI.
\end{theorem}

@@

\begin{theorem}[Weak isomorphism preserves TINI]
  \label{thm:iso-tini}
  If a language $L$ is weakly isomorphic to a language $L'$, and $L'$
  satisfies TINI, then $L$ satisfies TINI.
\end{theorem}

@@

\begin{definition}[Restricted IFC language]
  \label{def:restricted}
  For a family of predicates $\mathcal P$ (one for every reduction
  rule), we call
  |restrictedLang alpha targetLang| a restricted IFC language
  if its definition is equivalent to the abstract language
  |specLang alpha targetLang|, with the following exception:
  the reduction rules are restricted
  by adding a predicate $P \in \mathcal P$ to the premise of
  all rules other than \textsc{I-noStep}.  Furthermore, the predicate $P$
  can depend only on the \textit{erased} configuration
  |erase il ic|, where |il| is the label of the first task
  in the task list and |ic| is the full configuration.
\end{definition}

### 2015-01-15 paper/extensions.tex
presentational issue from reviewer 1
pass over everything except related and conclusion

While the IFC language presented thus far provides the basic
information flow primitives, actual IFC implementations
may wish to extend the minimal system with more specialized
constructs.
For example, COWL provides a labeled version of the XHR object.
%while LIO implements a labeled file system, used in server-side web
%applications~\cite{hails}.
%
Our system can be extended with constructs such as labeled values,
labeled mutable references, clearance, and privileges.
For space reasons, we provide
details of this, including the soundness
proof with the extensions, in \appendixextfirst{}.
%
Here, we instead discuss a limitation of our formalism: the lack of
external effects.

@@

Like labeled references, features with external effects must be
modeled in the IFC language; we must reason about the precise security
implications of features that otherwise inherently leak data.
%
Previous approaches have modeled external effects by internalizing the
effects as writes/reads to labeled channels/references~\cite{stefan:addressing-covert}.
%
Alternatively, it is possible to model such effects as messages to/from
certain labeled tasks, an approach taken by our Node.js
implementation.
%
These ``special'' tasks are trusted with access to the unlabeled
primitives that can be used to perform the external effects; since the
interface to these tasks is already part of the IFC language, the
proof only requires showing that this task does not leak information.
%
Instead of restricting or wrapping unsafe primitives,
COWL allow for
controlled network communication at the context boundary.
%
(By restricting the default XHR object, for example, COWL allows code
to communicate with hosts according to the task's current label.)

>>>>

While the IFC language presented thus far provides the basic
information flow primitives, actual IFC implementations
may wish to extend the minimal system with more specialized
constructs.
For example, COWL provides a labeled version of the XMLHttpRequest (XHR) object,
which is used to make network requests.
%while LIO implements a labeled file system, used in server-side web
%applications~\cite{hails}.
%
Our system can be extended with constructs such as labeled values,
labeled mutable references, clearance, and privileges.
For space reasons, we provide
details of this, including the soundness
proof with the extensions, in \appendixextfirst{}.
%
Here, we instead discuss a limitation of our formalism: the lack of
external effects.

@@

Like labeled references, features with external effects must be
modeled in the IFC language; we must reason about the precise security
implications of features that otherwise inherently leak data.
%
Previous approaches have modeled external effects by internalizing the
effects as operations on labeled channels/references~\cite{stefan:addressing-covert}.
%
Alternatively, it is possible to model such effects as messages to/from
certain labeled tasks, an approach taken by our Node.js
implementation.
%
These ``special'' tasks are trusted with access to the unlabeled
primitives that can be used to perform the external effects; since the
interface to these tasks is already part of the IFC language, the
proof only requires showing that this task does not leak information.
%
Instead of restricting or wrapping unsafe primitives,
COWL allow for
controlled network communication at the context boundary.
%
(By restricting the default XHR object, for example, COWL allows code
to communicate with hosts according to the task's current label.)

### 2015-01-15 paper/real.tex
pass over everything except related and conclusion

Luckily, this issue is not unique to our work---browser layout engines
also rely on isolating code executing in separate iframes (e.g., according to the
same-origin policy).
%
Since creating an OS thread for each iframe is expensive, both
the V8 and SpiderMonkey JavaScript engines provide a means for running
JavaScript code in isolation in a single OS thread,
on disjoint sub-heaps.
%
In V8, this unit of isolation is called a \emph{context}; in
SpiderMonkey, it is called a \emph{compartment}.
%
(We will use these terms interchangeably.)
%
Each context is associated with a global object, which, by
default, implements the JavaScript standard library (e.g.,
\verb|Object|, \verb|Array|, etc.).
%
Naturally, we adopt contexts to implement our notion of tasks.

When JavaScript is embedded in browser layout engines,
or in server-side platforms such as Node.js,
additional APIs such as the Document Object Model (DOM) or the file
system get exposed as part of the runtime system.
These features are exposed by extending the global object, just like
the standard library.  For this reason, it is easy to modify
these systems to forbid external effects when implementing
an IFC system, so important effects can be reintroduced in a safe manner.

@@

Furthermore, COWL repurposes existing contexts (e.g., iframes and
pages) as IFC tasks, only imposing additional constraints on how they
communicate.
%
As with Node.js, at its core, the global object of a COWL task
should only contain the standard JavaScript libraries and
\texttt{postMessage}, whose semantics are modeled by our
\textsc{JS-send} rule.
%
However, existing contexts have objects such as the DOM, which
demands COWL to restrict a task's external effects.
%
To this end, COWL mediates any communication (even via the DOM) at
the context boundary. %by COWL.

Simply disallowing all the external effects is overly-restricting for
real-world applications (e.g., pages typically load images, perform
network requests, etc.).
%
In this light, COWL allows safe network communication by associating an
implicit label with remote hosts (a host's label corresponds to
its origin).
%
In turn, when a task perform a request, COWL's IFC monitor
ensures that the task label can flow to the remote origin label.
%
While the external effects of COWL can be formally modeled, we do not
model them in our formalism, %more general framework,
since, like for the
Node.js case, they are specific to this system.
%
%Indeed, even the labeled HTTP client in  Node.js and COWL differ, even
%though they're both JavaScript systems, simply because of the
%differing environments.

@@

%Indeed, by rephrasing LIO in terms of our framework, we simplified
%LIO's treatment of exceptions.
%
Unlike our proposal, LIO originally associated labels with exceptions, in a
similar style to fine-grained
systems~\cite{stefan:2012:arxiv-flexible, Hritcu:2013:YIB:2497621.2498098}.
%
In addition to being overly complex, the interaction of exceptions
with clearance (which sets an upper bound on the floating label, see
Section~\ref{sec:clearance}) was incorrect: the clearance
was restored to the clearance at point of the catch.
%
Furthermore, pure exceptions (e.g., divide by zero) always percolated to
trusted code, effectively allowing for denial of service attacks.
%
The insights gained when viewing coarse-grained IFC as presented in this
paper led to a much cleaner, simpler treatment of exceptions in LIO,
which has now been adopted by LIO.
%By using isolated tasks, and thus treating the current label (and
%clearance) as global, the underlying Haskell exception system can be
%used directly: exceptional control flow does not affect IFC.

\subsection{C}
\label{sec:real:c}
%
C programs are able to execute arbitrary (machine) code, access
arbitrary memory, and perform arbitrary system calls.
%
Thus, the confinement of C programs must be imposed by the underlying OS
and hardware.
%
For instance, our notion of isolation can be achieved using Dune's
hardware protection mechanisms~\cite{Belay:2012:DSU:2387880.2387913},
similar to a simpler version of
Wedge~\cite{Belay:2012:DSU:2387880.2387913,
Bittau:2008:WSA:1387589.1387611} with an information flow control
policy.
%
Using page tables, a (trusted) IFC runtime could ensure that each task,
implemented as a lightweight process, can only access the memory it
allocates---tasks do not have access to any shared memory.
%
In addition, ring protection could be used to intercept system
calls performed by
a task and only permit those corresponding to our IFC language (such as
|getLabel| or |send|).
%
Dune's hardware protection mechanism would allow us to provide a concrete
implementation that is efficient and relatively simple to reason
about, but other sandboxing mechanisms could be used in place of Dune.

In this setting, the combined language of Section~\ref{sec:retrofit}
can be interpreted in the following way: calling from the target
language to the IFC language corresponds to invoking a system call.
%
Creating a new task with the |sandbox| system call corresponds to
\emph{forking} a process.  Using page tables, we can ensure that
there will be no shared memory
(effectively
defining |ap klone tS
= tS0|, where |tS0| is the set of pages necessary to bootstrap a
lightweight process).
%
Similarly, control over page tables and protection bits allows us to
define a |send| syscall that copies pages to our
(trusted) runtime queue; and, correspondingly, a |recv| that copies
the pages from the runtime queue to the (untrusted) receiver.
%
Since C is not memory safe, conditions on these system calls are
meaningless.
%
We leave the implementation of this IFC system for C as future work.

>>>>

Luckily, this issue is not unique to our work---browser layout engines
also rely on isolating code executing in separate iframes (e.g., according to the
same-origin policy).
%
Since creating an OS thread for each iframe is expensive, both
the V8 and SpiderMonkey JavaScript engines provide means for running
JavaScript code in isolation within a single OS thread,
on disjoint sub-heaps.
%
In V8, this unit of isolation is called a \emph{context}; in
SpiderMonkey, it is called a \emph{compartment}.
%
(We will use these terms interchangeably.)
%
Each context is associated with a global object, which, by
default, implements the JavaScript standard library (e.g.,
\verb|Object|, \verb|Array|, etc.).
%
Naturally, we adopt contexts to implement our notion of tasks.

When JavaScript is embedded in browser layout engines,
or in server-side platforms such as Node.js,
additional APIs such as the Document Object Model (DOM) or the file
system get exposed as part of the runtime system.
These features are exposed by extending the global object, just like
the standard library.  For this reason, it is easy to modify
these systems to forbid external effects when implementing
an IFC system, ensuring that important effects can be reintroduced in a safe manner.

@@

Furthermore, COWL repurposes existing contexts (e.g., iframes and
pages) as IFC tasks, only imposing additional constraints on how they
communicate.
%
As with Node.js, at its core, the global object of a COWL task
should only contain the standard JavaScript libraries and
\texttt{postMessage}, whose semantics are modeled by our
\textsc{JS-send} rule.
%
However, existing contexts have objects such as the DOM, which
require COWL to restrict a task's external effects.
%
To this end, COWL mediates any communication (even via the DOM) at
the context boundary. %by COWL.

Simply disallowing all the external effects is overly-restricting for
real-world applications (e.g., pages typically load images, perform
network requests, etc.).
%
In this light, COWL allows safe network communication by associating an
implicit label with remote hosts (a host's label corresponds to
its origin).
%
In turn, when a task performs a request, COWL's IFC monitor
ensures that the task label can flow to the remote origin label.
%
While the external effects of COWL can be formally modeled, we do not
model them in our formalism, %more general framework,
since, like for the
Node.js case, they are specific to this system.
%
%Indeed, even the labeled HTTP client in  Node.js and COWL differ, even
%though they're both JavaScript systems, simply because of the
%differing environments.

@@

%Indeed, by rephrasing LIO in terms of our framework, we simplified
%LIO's treatment of exceptions.
%
Unlike our proposal, LIO originally associated labels with exceptions, in a
similar style to fine-grained
systems~\cite{stefan:2012:arxiv-flexible, Hritcu:2013:YIB:2497621.2498098}.
%
In addition to being overly complex, the interaction of exceptions
with clearance (which sets an upper bound on the floating label, see
Section~\ref{sec:clearance}) was incorrect: the clearance
was restored to the clearance at point of the catch.
%
Furthermore, pure exceptions (e.g., divide by zero) always percolated to
trusted code, effectively allowing for denial of service attacks.
%
The insights gained when viewing coarse-grained IFC as presented in this
paper led to a much cleaner, simpler treatment of exceptions,
which has now been adopted by LIO.
%By using isolated tasks, and thus treating the current label (and
%clearance) as global, the underlying Haskell exception system can be
%used directly: exceptional control flow does not affect IFC.

\subsection{C}
\label{sec:real:c}
%
C programs are able to execute arbitrary (machine) code, access
arbitrary memory, and perform arbitrary system calls.
%
Thus, the confinement of C programs must be imposed by the underlying OS
and hardware.
%
For instance, our notion of isolation can be achieved using Dune's
hardware protection mechanisms~\cite{Belay:2012:DSU:2387880.2387913},
similar to 
Wedge~\cite{Belay:2012:DSU:2387880.2387913,
Bittau:2008:WSA:1387589.1387611}, but using an information flow control
policy.
%
Using page tables, a (trusted) IFC runtime could ensure that each task,
implemented as a lightweight process, can only access the memory it
allocates---tasks do not have access to any shared memory.
%
In addition, ring protection could be used to intercept system
calls performed by
a task and only permit those corresponding to our IFC language (such as
|getLabel| or |send|).
%
Dune's hardware protection mechanism would allow us to provide a concrete
implementation that is efficient and relatively simple to reason
about, but other sandboxing mechanisms could be used in place of Dune.

In this setting, the combined language of Section~\ref{sec:retrofit}
can be interpreted in the following way: calling from the target
language to the IFC language corresponds to invoking a system call.
%
Creating a new task with the |sandbox| system call corresponds to
\emph{forking} a process.  Using page tables, we can ensure that
there will be no shared memory
(effectively
defining |ap klone tS
= tS0|, where |tS0| is the set of pages necessary to bootstrap a
lightweight process).
%
Similarly, control over page tables and protection bits allows us to
define a |send| system call that copies pages to our
(trusted) runtime queue; and, correspondingly, a |recv| that copies
the pages from the runtime queue to the (untrusted) receiver.
%
Since C is not memory safe, conditions on these system calls are
meaningless.
%
We leave the implementation of this IFC system for C as future work.

### 2015-01-15 paper/retrofit.tex
pass over everything except related and conclusion

Information flow control systems operate by associating data with \emph{labels},
and specifying whether or not data tagged with one label |l1| can flow
to another
label |l2| (written as |canFlowTo2 l1 l2|).  These labels encode the desired
security policy (for example, confidential information should not flow to
a public channel), and the work of specifying the semantics of an information
flow language involves demonstrating that impermissible flows cannot happen,
a property called \emph{non-interference}.
%
In our coarse-grained floating-label approach, labels are associated with tasks.
%
The task label---we refer to the label of the currently executing task as the
\emph{current label}---serves to protect everything in the task's scope;
all data in a task shares this common label.

@@

Our goal now is to
describe how to take a \textbf{{\color{red} target
language}} with a formal operational semantics and combine it with an
\textit{{\color{blue} information flow control language}}.  For example,
taking ECMAScript as the target language and combining it with our IFC
language should produce the formal semantics for the core part of COWL~\cite{swapi}.  In this
presentation, we use a simple, untyped lambda calculus with mutable
references and fixpoint in place of ECMAScript, to demonstrate some of the key
properties without including too much detail (and because the embedding really
does not care about what, exactly, is in the target language); we discuss the proper
embedding in more detail in Section~\ref{sec:real}.

\vspace{1pt}
\noindent
\textit{Notation}
We have typeset nonterminals of the target language using \textbf{{\color{red}
bold font}} while the nonterminals of the IFC language have been typeset
with \textit{{\color{blue} italic font}}.  Readers are encouraged to view
a color copy of this paper, where target language nonterminals are colored \textbf{{\color{red} red}}
while IFC language nonterminals are colored \textit{{\color{blue} blue}}.

@@

These operations are all defined with respect to an evaluation context
|curlyBlueE| that represents the context of the current task which.
The evaluation context has three important pieces of
state: the global message queues |iS|, the current label |il| and the task ID |id|.

>>>>

Information flow control systems operate by associating data with \emph{labels},
and specifying whether or not data tagged with one label |l1| can flow
to another
label |l2| (written as |canFlowTo2 l1 l2|).  These labels encode the desired
security policy (for example, confidential information should not flow to
a public channel), while the work of specifying the semantics of an information
flow language involves demonstrating that impermissible flows cannot happen,
a property called \emph{non-interference}~\cite{Goguen82}.
%
In our coarse-grained floating-label approach, labels are associated with tasks.
%
The task label---we refer to the label of the currently executing task as the
\emph{current label}---serves to protect everything in the task's scope;
all data in a task shares this common label.

@@

Our goal now is to
describe how to take a \textbf{{\color{red} target
language}} with a formal operational semantics and combine it with an
\textit{{\color{blue} information flow control language}}.  For example,
taking ECMAScript as the target language and combining it with our IFC
language should produce the formal semantics for the core part of COWL~\cite{swapi}.  In this
presentation, we use a simple, untyped lambda calculus with mutable
references and fixpoint in place of ECMAScript to demonstrate some the key
properties of the system (and, because the embedding does not care
about the target language features); we discuss the proper embedding
in more detail in Section~\ref{sec:real}.

\vspace{1pt}
\noindent
\textit{Notation}
We have typeset nonterminals of the target language using \textbf{{\color{red}
bold font}} while the nonterminals of the IFC language have been typeset
with \textit{{\color{blue} italic font}}.  Readers are encouraged to view
a color copy of this paper, where target language nonterminals are colored \textbf{{\color{red} red}}
and IFC language nonterminals are colored \textit{{\color{blue} blue}}.

@@

These operations are all defined with respect to an evaluation context
|curlyBlueE| that represents the context of the current task.
The evaluation context has three important pieces of
state: the global message queues |iS|, the current label |il| and the task ID |id|.
